Musical elements in the discrete-time representation of sound

### 2017-08-08 article/acm/body.tex
finished all notes before last read

\section{Organization of notes in music}\label{notasMusica} \label{sec:notesMusic}
Consider $S_j=\left\{  s_j=T_i^j=\{t_i^j\}_{i=0}^{\Lambda_j-1} \right\}_{j=0}^{H-1}$, the sequence $S_j$ of $H$ musical events $s_j$. Let $S_j$ be a `musical structure', composed by events $s_j$ which are musical structures themselves, e.g.\ notes. This section is dedicated to techniques that make $S_j$ interesting and enjoyable for audition. More specifically, what follows is a summary of academic music composition theory and praxis.
This section does not benefit from equations that dictate the amplitude of each sample as deeply as the previous sections.
Even so, we understand that this content is the missing piece for synthesizing music given all the content above.
The concepts are given algorithmic implementations in \mass~\cite{SI}.

@@

% scramble 
From these processes, many others are derived, such as the inverted retrograde, the temporal contraction with an external suffix, etc. Variations are often thought about in the terms above but are also often very loose, such as an arbitrary shuffle of the notes in a melody which the composer finds interesting. As a result, a whole process of mental and neurological activity is unleashed for relating the arcs, responsible for feelings, memories and imaginations, typical of a diligent music listening. This cortical activity is critical to
musical therapy, known by its utility in cases of depression and neurological injury. Also, tt is known that regions of the human brain responsible for sonic processing are also used for other activities, such as for performing verbal discourse and mathematics~\cite{Sacks,Roederer}.

@@

@@

\subsection{Serialism and post-serial techniques}
Expanding the concepts in Section~\ref{sec:atonal},
sequences of characteristics can be predefined and
used throughout a musical piece.
These sequences can be of intensities, timbre, durations,
density of events, etc.
The sequences can be used very strictly or loose,
such as by skipping some elements.
The sequences can be of different sizes, yielding arcs
until the initial condition is reached again (i.e. cycles).
One paradigmatic case is the ``total serialism'' where all
the musical characteristics are serialized.
Although the use of sequences is inherent to music (e.g. scales, metric pulses),
their use with greater emphasis than tonal (or modal) elements
in western music, as an artistic trend, took place only in the
first half of the twentieth century and is called ``serialism''.
Post-serial techniques are numerous, but here is a description of
important concepts:
\begin{itemize}
	\item Spectralism:
	\item Spectromorphology:
	\item Stochastic music:
	\item Textures:
\end{itemize}

>>>>

\section{Organization of notes in music}\label{notasMusica} \label{sec:notesMusic}
Consider $S_j=\left\{  s_j=T_i^j=\{t_i^j\}_{i=0}^{\Lambda_j-1} \right\}_{j=0}^{H-1}$, the sequence $S_j$ of $H$ musical events $s_j$. Let $S_j$ be a `musical structure', composed by events $s_j$ which are musical structures themselves, e.g.\ notes. This section is dedicated to techniques that make $S_j$ interesting and enjoyable for audition. More specifically, what follows is a summary of academic music composition theory and praxis.
This section does not benefit from equations that dictate the amplitude of each sample as deeply as the previous sections.
Even so, we understand that this content is very useful for synthesizing music given all the content above.
The concepts are given algorithmic implementations in \mass~\cite{SI} and can be further formalized~\cite{topos},
although at a cost of intelligibility which we chose to avoid in this exposition.

@@

From these processes, many others are derived, such as the inverted retrograde, the temporal contraction with an external suffix, etc. Variations are often thought about in the terms above but are also often very loose, such as an arbitrary shuffle of the notes in a melody which the composer finds interesting. As a result, a whole process of mental and neurological activity is unleashed for relating the arcs, responsible for feelings, memories and imaginations, typical of a diligent music listening. This cortical activity is critical to
musical therapy, known by its utility in cases of depression and neurological injury. Also, it is known that regions of the human brain responsible for sonic processing are also used for other activities, such as for performing verbal discourse and mathematics~\cite{Sacks,Roederer}.

@@

For a scholastic discussion on the composition of motives, phrases, melodies, themes and music form (such as rondo, ternary, theme and variations) see~\cite{Schoenberg}.
% melody, answer and reply, motives, etc

@@

\subsection{Serialism and post-serial techniques}
Expanding the concepts in Section~\ref{sec:atonal},
sequences of characteristics can be predefined and
used throughout a musical piece.
These sequences can be of intensities, timbre, durations,
density of events, etc.
The sequences can be used very strictly or loose,
such as by skipping some elements.
The sequences can be of different sizes, yielding arcs
until the initial condition is reached again (i.e. cycles).
One paradigmatic case is the ``total serialism'' where all
the musical characteristics are serialized.
Although the use of sequences is inherent to music (e.g. scales, metric pulses),
their use with greater emphasis than tonal (or modal) elements
in western music, as an artistic trend, took place only in the
first half of the twentieth century and is called ``serialism''.
Post-serial techniques are numerous, but here is a description of
important concepts:
\begin{itemize}
	\item Spectralism:
		consists on use the (Fourier) spectrum of a sound or the harmonic series for music composition,
		such as
		to derive harmonies, sequences of pitches or a temporal evolution of the overall spectrum~\cite{grisey}.
		In other words, the most prominent frequencies can be used as pitches,
		real notes can be used to mimic an original spectrum
		(e.g. use piano notes to mimic the spectrum of a spoken sentence) or portions of the spectrum made to vary~\cite{spectral}.
	\item Spectromorphology:
		can be considered a spectral music (spectralism) theoretical framework~\cite{smalley,schaeffer}
		that examines the relation between sound spectra and their temporal evolution.
		The theory poses e.g. different onsets, continuations and terminations; characteristics of (sonic) ``motion'';
		and spectral density.
	\item Stochastic music:
		the use of random variables to describe musical elements are largely considered in stochastic music~\cite{formalized}.
		In summary, one can use probability distributions for the synthesis of sounds and larger scale musical elements.
		Changes in these distributions or in other characteristics yield the discourse.
	\item Textures:
		sounds can be assembled in terms of a ``sonic texture''.
		Here, a texture is often thought about very abstractly
		as a sonic counterpart of visual texture.
		Parameters that can be used for such are: range between highest and lowest note,
		density of notes, durations of notes, motives, number of voices, etc.
		If the sounds are small enough (typically < $100ms$) the process can be though of in terms of \emph{granular synthesis}~\cite{microsound}.
\end{itemize}
